---
title: "OpenIntro Lab - Chapter 1"
output:
  html_document:
    df_print: paged
---

# OpenIntro Lab - Introduction to data #

This notebook is the first chapter of my personal study book. The general aim is to learn how summarize and describe the raw information - the data. It is part of the first set of **OpenIntro** labs, that are part of the study of the book: **Introductory Statistics with Randomization and Simulation**. Lab avaliable from: [Introducton to Data](https://htmlpreview.github.io/?https://github.com/andrewpbray/oiLabs-base-R/blob/master/intro_to_data/intro_to_data.html)

We will use a data set collected by the **Centers for Disease Control and Prevention** (CDC), called **The Behavioral Risk Factor Surveillance System** (BRFSS). Avaliable from: [BRFSS](http://www.cdc.gov/brfss).

This R Notebooke is organized in the following way: there will always be a question about the dataset, followed by code chunks (R script) and its outputs. For some answers the output will be considered sufficient. For others a short answer will be written.

We will focus on a random sample of 20,000 people from the **BRFSS** survey conducted in 2000. While there are over 200 variables in this data set, we will work with a small subset. 

```{r}
source("http://www.openintro.org/stat/data/cdc.R")
```

### **Exercise 1** ### 
How many cases are there in this data set? How many variables? For each variable, identify its data type (e.g. categorical, discrete).

```{r}
names(cdc)
```

```{r}
head(cdc)
```

```{r}
tail(cdc)
```

```{r}
str(cdc)
```

**Answer**: 
There are 20000 cases and 9 variables in this dataset. We can see the types of each variable in the output above.


### **Exercise 2** ### 
Create a numerical summary for `height` and `age`, and compute the interquartile range for each. Compute the relative frequency distribution for `gender` and `exerany`. How many males are in the sample? What proportion of the sample reports being in excellent health?

```{r}
summary(cdc$height)
```

```{r, error=TRUE}
# interquartile range = 3rd Qu - 1st Qu
IQR_height <- 70-64
```

```{r, error=TRUE}
IQR_height
```

```{r}
summary(cdc$age)
```

```{r}
# interquartile range = 3rd Qu - 1st Qu
IQR_age <- (57-31)
```

```{r}
IQR_age
```

```{r}
table(cdc$gender)
```

```{r}
table(cdc$gender)/20000
```

```{r}
table(cdc$exerany)/20000
```

**Answer**: There are 9569 males. The frequency distribution for `gender` and `exerany`, can be seen int the output above.

```{r}
summary(cdc$genhlth) / 20000
```

**Answer**: the proportion of the sample reports being in excellent health = 0.23285 or 23%.

### **Exercise 3** ### 
What does the mosaic plot reveal about smoking habits and gender?

```{r}
gender_smoke <- table(cdc$gender,cdc$smoke100)
```

```{r}
gender_smoke
```

```{r}
mosaicplot(gender_smoke) #  column labels: 0 and 1. 1 indicates a respondent has smoked at least 100 cigarettes.
```

**Answer**: It seems that men smoke more often than women.
